Filmchat

### 2017-06-15 paper/paper.tex
Updates updates updates

\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{google_assistant.png}
  \caption{Google Assitant recommending a movie based on geographic context}
  \label{fig:google_assistant}
\end{figure}

@@

\section{Speech Recognizer}
The Speech recognizer, is based on the Google Speech API \cite{GoogleCloudPlatformSpeechPlatform} and is programmed in Csharp. The code is mostly just adapted to post to a web interface, in addition to printing to the console / cli. The Speech Recognizer listens to the detected microphone on the computer by default. It keeps a python dictionary of the users, and their, movie and their most recent sentiment score in memory.

\section{Omdbsearcher}
The Omdbsearcher is based on the Open Movie Data base API.\cite{OMDbDatabase}. When it gets text it will try to search for this text on the OMDB Api and return the JSON formatted result if it finds something, or return a negative result if it does not find anything. The web service listens to port 3000 by default.

@@

The last input, would have used the same modules, but would have gotten their input from a chat room. The difference would have been the more asynchronous behavior of that kind of interaction, but sentiments and information could have been harvested in much the same way.
\section{Design rationale}
In the Brave NUI World\cite{Wigdor2010BraveWorld} Wigdor and Wilson leads by saying the following about Natural User interfaces and their design guidelines:
\begin{itemize}
\item Create an experience that, for expert users, can feel like an extension of their body.
\item Create an experience that feels just as natural to a novice as it does to an
expert user.
\item Create an experience that is authentic to the medium—do not start by trying to
mimic the real world or anything else.
\item Build a user interface that considers context, including the right metaphors,
visual indications, feedback, and input/output methods for the context.
\item Avoid falling into the trap of copying existing user interface paradigms.
\end{itemize}

@@

\section{Cost / Benefit of selected interaction techniques}
\subsection{Chat}
Chat is in my view has become a natural way of interacting, several times a day, I interact through chat, be it privately and also work related. A non-exhaustive list of some of the chat interaction I use during a day, or interactions close to chat, are:
\begin{itemize}
\item SMS or Textmessaging
\item Microsoft Teams
\item Slack
\item Lync for Business (formerly Skype)
\item Facebook Chat
\item WhatsApp
\item Snapchat 
\item Tinder
\item Happn
\item Chat Support on company web pages
\end{itemize}

@@

First I tried using the Node JS library, but I was not able to get the sound recorder to work properly on my Windows 10 laptop. (It might work better on a linux or Mac, but I am not sure of that)

@@

I then created a local web server serving the incoming recognized speech, and then trying to search for them on OMDB. If it got a hit, it posted it to the Slack Channel I had selected.
\newpage
\subsection{Omdbsearcher listener route}
\begin{verbatim}
app.post("/", function(req, res) {
    if (req.body.text) {
      console.log("Searching for on OMDB: " + req.body.text)
        options.qs.t = req.body.text;
        request(options, function(error, response, body) {
            if (error) throw new Error(error);
            postToSlack(JSON.parse(body), options.qs.t)
            res.send("OK");
        });
     } else {
     res.send("Got no input");
    }
});
\end{verbatim}
This listener route listens to a HTTP POST JSON with "text" key. It then searches for the text in OMDB. Either way it posts to the Slackchannel using the postToSlack function.
\newpage
\subsection{Omdbsearcher post to slack}
\begin{verbatim}
function postToSlack(result, searchterm) {
    var slackMessage;
    if (result.Response != "True") {
        console.log(result);
        slackMessage = "Movie with searchterm: " + searchterm + " not found!";
      } else {
        slackMessage = formatResult(result);
    }
    var slackOptions = {
        method: 'POST',
        url: config.slackurl,
        headers: {
            'cache-control': 'no-cache',
            'content-type': 'application/json',
            accept: 'application/json'
        },
        body: slackMessage,
        json: true
    };
    request(slackOptions, function(error, response, body) {
        if (error) throw new Error(error);
        console.log(body);
    });
}
\end{verbatim}
This function posts the formatted result to Slack. To be able to do this you have to generate the key which is hardcoded into the "config.slackurl" string.

\section{Textblob Botlistener}
I then checked out a couple of different ways of interpreting language. Among them I looked at Wit.AI\cite{Wit.aiWit.ai}, and also looked at API.AI\cite{API.AIConversationalAPI.AI}. While both of them looked promising, the scope of this project meant I should probably use something simpler and faster to implement. 
From other projects, I already have the knowledge that Python has such libraries that are both good, and with good performance. I therefore winded up with TextBlob, which is a library that can both get sentiment, and noun phrases\cite{WikipediaNounPhrase} from sentences. \cite{TextBlobTutorial:Documentation} 
I therefore just had to create a web service interface for the TextBlob, and was able to do this quickly.
\newpage
\begin{verbatim}
@app.route('/api/', methods=['POST'])
def api_post():
    #text = request.json['text']
    text = request.form['text']
    #username = request.json['user_name']
    username = request.form['user_name']
    if (username == "slackbot"):
        return jsonify("OK")
    else:
        testimonial = TextBlob(text)
        sent = testimonial.sentiment.polarity
        print (testimonial.noun_phrases)
        if (len(testimonial.noun_phrases) > 0):
            noun = testimonial.noun_phrases[0]
        else:
            noun = "Did not find movie"
        findMovieTitle(noun)
        if (username in users):
            users[username][noun] = sent
        else:
            users[username] = {noun: sent}
        print (users) #Debug
        return jsonify(text = username + " has a sentiment value " + str(sent) + " for " + noun)
\end{verbatim}
This code snippet listens to the route "/api/" it gets the text and user name formatted as application/x-www-form-urlencoded and then using the TextBlob library extracts the sentiment\cite{WikipediaSentimentAnalysis} and noun phrase\cite{WikipediaNounPhrase}.

\section{Spelling corrector}
While the Textblob Botlistener was able to get the noun phrase\cite{WikipediaNounPhrase} of sentences about movies, and that this sometimes is an actual movie, it is not always the case. Therefore I thought about using some auto-correct or spelling corrections, but using the freely available IMDB datasets as valid words \cite{IMDbAlternativeInterfaces}. I found a good example by Peter Norvig \cite{NorvigHowCorrector} on how to do it.
But both the development, and the performance was rather slow, and the dataset would have needed quite a lot of cleaning before I could have used it. Therefore I decided on not to implement this part, and to put it off to a possible future work. The beginning work can be seen in the Github\cite{Github2017GitHubProject} under the did-you-mean folder.

@@

\chapterimage{head3.png}
\chapter{Ten Meaningful Screens}
\section{Speech recognition}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{speech_recognition_c-sharp.PNG}
  \caption{Speech recognition using the CSharp SDK for Google Speech}
  \label{fig:c-sharp-speech}
\end{figure}
To get to this screen Fig\ref{fig:c-sharp-speech}, you have to compile the C\# project located under speech-recognizer folder in the GitHub project. \cite{Github2017GitHubProject}
Then you would have to enable the Speech API in the Google Cloud Console, and then finally compile the C\# and run it.
When you are able to execute the program, you type in "Recognize.exe listen". The program will then, using the microphone it detects on your system. Listen for audio and send this to the Speech API for recognition. It will then return them to the cli, and also post them to the Omdbsearcher.

\newpage
\section{Using the Omdbsearcher}
The omdbsearcher Fig\ref{fig:omdbsearcher} part of Filmchat is located under the Omdsearcher folder in the GitHub project \cite{Github2017GitHubProject}. It is based on NodeJS, so first step is to install the neccessary packages by doing: \"npm install\". Then you start it by executing the command: \"node index.js\". This will in turn start a small webserver running on port 3000.
It listens for JSON-formatted data.
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{omdbsearcher.PNG}
  \caption{Omdbsearcher in the cli}
  \label{fig:omdbsearcher}
\end{figure}
For debugging purposes it outputs the search results from the JSON POST it got.

@@

\newpage
\section{Getting the status from the TextBlob Botlistener}
To get the current status of the users and their sentiments around movies, a simple GET to the /api/ will return a JSON formatted result as seen in Fig\ref{fig:sending-botlistener} Negative values are negative sentiments, and as closer the value gets to 1, the better the sentiment.
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{get_textblob.PNG}
  \caption{Getting status of sentiments}
  \label{fig:sentiment-status}
\end{figure}

\newpage
\section{Everything together, almost}
Here \(Fig\ref{fig:everything-together}\), almost all the components are playing together. I write a chat message in the designated Slack channel, while running both the Omdbsearcher, and the TextBlob botlistener, and the result, along with the sentiment gets posted to the channel.
The only part, not here, is the Speech Recognizer, as I was not able to make it recognize more than a word, here and there.

@@

@@

\chapterimage{head3.png}
\chapter{Two Valuable Outcomes}
Generally speaking I would say that, unfortunately my interactive system does not provide that valuable of outcomes.
However here I will try to describe the two, that I thought came the closest.

\section{Speech recognition of movies}
Speech recognition of movies. Being able to search for, and find ratings for movies without having to use the telephone, could be something I might use.
I also thik that this could be a short term valuable outcome of the system. Especially if there was a way of starting to play the movie automatically, once I was happy with the selection. However for a home theater version of this to work, the description and possibly the ability to play a trailer would have to be done on the Television, and maybe not in a Slack channel.

>>>>

\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{google_assistant.png}
  \caption{Google Assistant recommending a movie based on geographic context}
  \label{fig:google_assistant}
\end{figure}

@@

\section{Speech Recognizer}
The Speech recognizer, is based on the Google Speech API \cite{GoogleCloudPlatformSpeechPlatform} and is programmed in C sharp. The code is mostly just adapted to post to a web interface, in addition to printing to the console / cli. The Speech Recognizer listens to the detected microphone on the computer by default. It keeps a python dictionary of the users, and their, movie and their most recent sentiment score in memory.

\section{Omdbsearcher}
The Omdbsearcher is based on the Open Movie Data base API.\cite{OMDbDatabase}. When it gets text it will try to search for this text on the OMDB API and return the JSON formatted result if it finds something, or return a negative result if it does not find anything. The web service listens to port 3000 by default.

@@

The last input, would have used the same modules, but would have gotten their input from a chat room. The difference would have been the more asynchronous behavior of that kind of interaction, but sentiments and information could have been harvested in much the same way.
\section{Design rationale}
In the Brave NUI World\cite{Wigdor2010BraveWorld} Wigdor and Wilson leads by saying the following about Natural User interfaces and their design guidelines:
\begin{itemize}
\item Create an experience that, for expert users, can feel like an extension of their body.
\item Create an experience that feels just as natural to a novice as it does to an
expert user.
\item Create an experience that is authentic to the medium, do not start by trying to
mimic the real world or anything else.
\item Build a user interface that considers context, including the right metaphors,
visual indications, feedback, and input/output methods for the context.
\item Avoid falling into the trap of copying existing user interface paradigms.
\end{itemize}

@@

\section{Cost / Benefit of selected interaction techniques}
\subsection{Chat}
Chat is in my view has become a natural way of interacting, several times a day, I interact through chat, be it privately and also work related. A non-exhaustive list of some of the chat interaction I use during a day, or interactions close to chat, are:
\begin{itemize}
\item SMS or Text messaging
\item Microsoft Teams
\item Slack
\item Lync for Business (formerly Skype)
\item Facebook Chat
\item WhatsApp
\item Snapchat 
\item Tinder
\item Happn
\item Chat Support on company web pages
\end{itemize}

@@

First I tried using the Node JS library, but I was not able to get the sound recorder to work properly on my Windows 10 laptop. (It might work better on a Linux or Mac, but I am not sure of that)

@@

I then created a local web server serving the incoming recognized speech, and then trying to search for them on OMDB. If it got a hit, it posted it to the Slack Channel I had selected.
\newpage
\subsection{Omdbsearcher listener route}
\begin{verbatim}
app.post("/", function(req, res) {
    if (req.body.text) {
      console.log("Searching for on OMDB: " + req.body.text)
        options.qs.t = req.body.text;
        request(options, function(error, response, body) {
            if (error) throw new Error(error);
            postToSlack(JSON.parse(body), options.qs.t)
            res.send("OK");
        });
     } else {
     res.send("Got no input");
    }
});
\end{verbatim}
This listener route listens to a HTTP POST JSON with "text" key. It then searches for the text in OMDB. Either way it posts to the Slackchannel using the postToSlack function.
\newpage
\subsection{Omdbsearcher post to slack}
\begin{verbatim}
function postToSlack(result, searchterm) {
    var slackMessage;
    if (result.Response != "True") {
        console.log(result);
        slackMessage = "Movie with searchterm: " + searchterm + " not found!";
      } else {
        slackMessage = formatResult(result);
    }
    var slackOptions = {
        method: 'POST',
        url: config.slackurl,
        headers: {
            'cache-control': 'no-cache',
            'content-type': 'application/json',
            accept: 'application/json'
        },
        body: slackMessage,
        json: true
    };
    request(slackOptions, function(error, response, body) {
        if (error) throw new Error(error);
        console.log(body);
    });
}
\end{verbatim}
This function posts the formatted result to Slack. To be able to do this you have to generate the key which is hard coded into the "config.slackurl" string.

\section{TextBlob Botlistener}
I then checked out a couple of different ways of interpreting language. Among them I looked at Wit.AI\cite{Wit.aiWit.ai}, and also looked at API.AI\cite{API.AIConversationalAPI.AI}. While both of them looked promising, the scope of this project meant I should probably use something simpler and faster to implement. 
From other projects, I already have the knowledge that Python has such libraries that are both good, and with good performance. I therefore winded up with TextBlob, which is a library that can both get sentiment, and noun phrases\cite{WikipediaNounPhrase} from sentences. \cite{TextBlobTutorial:Documentation} 
I therefore just had to create a web service interface for the TextBlob, and was able to do this quickly.
\newpage
\begin{verbatim}
@app.route('/api/', methods=['POST'])
def api_post():
    #text = request.json['text']
    text = request.form['text']
    #username = request.json['user_name']
    username = request.form['user_name']
    if (username == "slackbot"):
        return jsonify("OK")
    else:
        testimonial = TextBlob(text)
        sent = testimonial.sentiment.polarity
        print (testimonial.noun_phrases)
        if (len(testimonial.noun_phrases) > 0):
            noun = testimonial.noun_phrases[0]
        else:
            noun = "Did not find movie"
        findMovieTitle(noun)
        if (username in users):
            users[username][noun] = sent
        else:
            users[username] = {noun: sent}
        print (users) #Debug
        return jsonify(text = username + " has a sentiment value " + str(sent) + " for " + noun)
\end{verbatim}
This code snippet listens to the route "/api/" it gets the text and user name formatted as application/x-www-form-urlencoded and then using the TextBlob library extracts the sentiment\cite{WikipediaSentimentAnalysis} and noun phrase\cite{WikipediaNounPhrase}.

\section{Spelling corrector}
While the TextBlob Botlistener was able to get the noun phrase\cite{WikipediaNounPhrase} of sentences about movies, and that this sometimes is an actual movie, it is not always the case. Therefore I thought about using some auto-correct or spelling corrections, but using the freely available IMDB datasets as valid words \cite{IMDbAlternativeInterfaces}. I found a good example by Peter Norvig \cite{NorvigHowCorrector} on how to do it.
But both the development, and the performance was rather slow, and the dataset would have needed quite a lot of cleaning before I could have used it. Therefore I decided on not to implement this part, and to put it off to a possible future work. The beginning work can be seen in the GitHub\cite{Github2017GitHubProject} under the did-you-mean folder.

@@

\chapterimage{head3.png}
\chapter{Ten Meaningful Screens}
\section{Speech recognition}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{speech_recognition_c-sharp.PNG}
  \caption{Speech recognition using the C Sharp SDK for Google Speech}
  \label{fig:c-sharp-speech}
\end{figure}
To get to this screen Fig\ref{fig:c-sharp-speech}, you have to compile the C\# project located under speech-recognizer folder in the GitHub project. \cite{Github2017GitHubProject}
Then you would have to enable the Speech API in the Google Cloud Console, and then finally compile the C\# and run it.
When you are able to execute the program, you type in "Recognize.exe listen". The program will then, using the microphone it detects on your system. Listen for audio and send this to the Speech API for recognition. It will then return them to the cli, and also post them to the Omdbsearcher.

\newpage
\section{Using the Omdbsearcher}
The Omdbsearcher Fig\ref{fig:omdbsearcher} part of Filmchat is located under the Omdbsearcher folder in the GitHub project \cite{Github2017GitHubProject}. It is based on NodeJS, so first step is to install the necessary packages by doing: \"npm install\". Then you start it by executing the command: \"node index.js\". This will in turn start a small web server running on port 3000.
It listens for JSON-formatted data.
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{omdbsearcher.PNG}
  \caption{Omdbsearcher in the cli}
  \label{fig:omdbsearcher}
\end{figure}
For debugging purposes it outputs the search results from the JSON POST it got.

@@

\newpage
\section{Getting the status from the TextBlob Botlistener}
To get the current status of the users and their sentiments around movies, a simple GET to the /api/ will return a JSON formatted result as seen in Fig\ref{fig:sending-botlistener} Negative values are negative sentiments, and as closer the value gets to 1, the better the sentiment. There is also the possibility of using a /sentiment command in the designated Slack channel. See Fig\ref{fig:sentiment-command}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{get_textblob.PNG}
  \caption{Getting status of sentiments}
  \label{fig:sentiment-status}
\end{figure}
\begin{figure}[]
  \centering
   \includegraphics[width=0.77\textwidth]{sentiments_command.PNG}
  \caption{Sentiment command}
  \label{fig:sentiment-command}
\end{figure}

\newpage
\section{Everything together, almost}
Here \(Fig\ref{fig:everything-together}\), almost all the components are playing together. I write a chat message in the designated Slack channel, while running both the Omdbsearcher, and the TextBlob Botlistener, and the result, along with the sentiment gets posted to the channel.
The only part, not here, is the Speech Recognizer, as I was not able to make it recognize more than a word, here and there.

@@

\section{User can access the status of the sentiments}
If a user in the Slack channel wants to check the current status of the sentiments, he or she can write /sentiments, and he will get the currents status from the TextBlob Botlistener.

@@

\chapterimage{head3.png}
\chapter{Two Valuable Outcomes}
Generally speaking I would say that, unfortunately my interactive system does not provide that valuable of outcomes.
However here I will try to describe the three, that I thought came the closest.

\section{Speech recognition of movies}
Speech recognition of movies. Being able to search for, and find ratings for movies without having to use the telephone, could be something I might use.
I also think that this could be a short term valuable outcome of the system. Especially if there was a way of starting to play the movie automatically, once I was happy with the selection. However for a home theater version of this to work, the description and possibly the ability to play a trailer would have to be done on the Television, and maybe not in a Slack channel.
